---
title: "Spatial Viz"
---

Use this file to generate a professional looking **spatial** visualization.  The visualization will not perfect the first time but you are expected to improve on it throughout the semester especially after covering advanced topics such as effective viz.

Create a data set using the `tribble()` function that has 10-15 rows of your favorite places. The columns will be the name of the location, the latitude, the longitude, and a column that indicates if it is in your top 3 favorite locations or not. For an example of how to use `tribble()`, look at the `favorite_stp` that was created manually below.

```{r}

# Load required libraries
library(clipr)
library(cpp11)
library(crayon)
library(data.table)
library(dbplyr)
library(dplyr)
library(dtplyr)
library(fivethirtyeight)
library(fivethirtyeightdata)
library(ggformula)
library(gtools)
library(hexbin)
library(labelled)
library(labeling)
library(leaflet)
library(leaflet.providers)
library(magrittr)
library(lubridate)
library(maps)
library(mosaic)
library(mosaicCore)
library(viridis)
library(tidyselect)
library(tidyr)
library(tidyverse)
library(ggrepel)
library()
```

```{r}
# David's Path in New Haven - Used Google Maps to get coordinates
# https://support.google.com/maps/answer/18539?hl=en&co=GENIE.Platform%3DDesktop
fave_yale <- tribble(
  ~place, ~lat, ~long, ~favorite,
  "Berkley College at Yale University", 41.31031346050375, -72.9281338032286, 2, 
  "Union Station in New Haven", 41.29768986562742, -72.92661307837868, 1,
  "Ashley's Ice Cream", 41.311430652674765, -72.92976153023119, 2,
  "Bingham Hall Old Campus", 41.30774207466993, -72.92879181859293, 1,  
  "Omni Hotel", 41.30580783013424, -72.92728978156289, 2,
  "Yale Science Hill", 41.31744466890471, -72.9218180751765, 1,
  "Beinecke Rare Books and Manuscripts Library", 41.31147136612109, -72.92738460510112, 1,
  "Yale Bookstore", 41.31210800597264, -72.9311289689019, 1
)
```

```{r}

```


```{r}
pal <- colorFactor(palette = c("red", "blue"), domain = fave_yale$favorite)
```


```{r}
#|fig-cap: "Map that showcases in blue and red the best spots on Yale University campus"
 #| fig-alt: "A map that plots the author's Top spots across Yale Univeristy"

title <- tags$d("My Map by David")
leaflet(data = fave_yale)|>
  addProviderTiles(providers$OpenStreetMap)|>
  addCircles(weight = 10, opacity = 0.8, color = ~pal(favorite), label = ~place)|>
  addLegend(position = "topright", pal = pal, values = ~favorite, title = "Best Spots at Yale University")|>
  addControl("<strong>David's Top Spots on Yale Campus</strong>", position = "bottomleft", className = "map-title") |>
  addControl("After a visit to Yale University Campus in New Haven, David highlights in blue the best spots to eat and in red the best spots to hangout. By David Manguluti, February 22 2025", position = "bottomleft", className = "map-caption")
```





