---
title: "Exam 1"
editor: visual
---
```{r}
 
 library(tidytuesdayR)
 library(tidyverse)
 
 tuesdata <- tt_load('2020-02-18')
 fc <- tuesdata$food_consumption
```

```{r}
 head(fc)
 dim(fc)
 str(fc)
```



```{r}
library(ggplot2)
library(dplyr)

fc %>%
  count(country) %>%
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of Entries per Country",
       x = "Country",
       y = "Count") +
  theme_minimal()

fc %>%
  count(food_category) %>%
  ggplot(aes(x = reorder(food_category, n), y = n)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Entries by Food Category",
       x = "Food Category",
       y = "Count") +
  theme_minimal()

ggplot(fc, aes(x = consumption)) +
  geom_histogram(binwidth = 10, fill = "tomato", color = "white") +
  labs(title = "Distribution of Food Consumption",
       x = "Consumption",
       y = "Frequency") +
  theme_minimal()
```

```{r}
 ggplot(fc, aes(x = country)) +
   geom_bar()

 ggplot(fc, aes(x = food_category)) +
   geom_bar()

 ggplot(fc, aes(y = consumption)) +
   geom_histogram()
```

Overall food consumption / food category
*This would be a simple bar chart*
Overall food consumption / country
*This would similarly be a simple bar chart*

For answering question 9, I would select a stacked bar chart.
  This chart would have axes labels, alt text, a caption, a legend, and a title.
  
Some further questions I have are
How does a country’s population size and land area influence its total and per capita food consumption?
(This invites exploration of both physical size and demographic scale.)
	2.	What is the relationship between a country’s CO₂ emissions and its patterns of food consumption?
(This allows for examining whether countries with higher emissions consume more resource-intensive foods.)
	3.	How do different food categories (e.g., meat, dairy, grains) contribute to CO₂ emissions across countries?
