---
title: "Mulivariate Viz"
---

Use this file for practice with the **mulivariate viz** in-class activity.  Refer to the class website for details.
#Pre-work
```{r}
library(tidyverse)

# Import data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

# Check out the first 6 rows
head(weather)
# What are the units of observation?
str(weather)

# How many data points do we have? 
nrow(weather)

# What type of variables do we have?
```
```{r}
ggplot(weather,aes(x=temp3pm))+
  geom_density()
```

```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density()
```
```{r}
# Plot 2 (no facets or densities)
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm)) + 
  geom_bar()
```
```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density()+
  facet_wrap(~ temp9am) 
```
```{r}
# Don't worry about the syntax (we'll learn it soon)
woll <- weather |>
  filter(location == "Wollongong") |> 
  mutate(date = as.Date(date))  
# How often does it raintoday?
# Fill your geometric layer with the color blue.
ggplot(woll, aes(x = raintoday))
```
```{r}
# If it does raintoday, what does this tell us about raintomorrow?
# Use your intuition first
# Now compare different approaches

# Default: stacked bars
# Plot temp3pm vs temp9am
# Change the code in order to indicate the location to which each data point corresponds
# Plot the relationship of raintomorrow & raintoday
# Change the code in order to indicate this relationship by location
ggplot(weather, aes(x = raintoday, fill = raintomorrow, alpha = location)) + 
  geom_bar(position = "fill")
  
```


#Exercise One
```{r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)
```
```{r}
ggplot(education,aes(x=sat))+
  geom_histogram()
```
Summarize your observations from the plot. Comment on the basics: range, typical outcomes, shape. (Any theories about what might explain this non-normal shape? 
-Range 1100 to zero
-The outlier presents an abnormality

#Exercise Two 
# Construct a plot of sat vs expend
# Include a "best fit linear regression model" (HINT: method = "lm")

```{r}
ggplot(education,aes(x=sat, y=expend))+
  geom_smooth(method = "lm")

# Construct a plot of sat vs salary
# Include a "best fit linear regression model" (HINT: method = "lm")

ggplot(education,aes(x=sat, y=salary))+
  geom_smooth(method = "lm")
```
Part b
What are the relationship trends between SAT scores and spending? Is there anything that surprises you?
There is an inverse relationship between spending, salaries and SAT score. This is very surprising as one would assume that more spending and higher wages would be proportionate to higher SAT scores. 


#Exercise Three
```{r}
ggplot(education, aes(x=sat,y=salary, color=expend))+
  geom_point(method = "lm")
```


#Exercise Four 
```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 3))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```


#Exercise Five
```{r}
ggplot(education, aes(x=fracCat))+
  geom_bar()
```
```{r}
ggplot(education, aes())
```


#Exercise Six



#Exercise Seven

# Load tidyverse package for plotting and wrangling
library(tidyverse)

```{r}
# Load tidyverse package for plotting and wrangling
library(tidyverse)

# Import the data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))

ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point()
```
```{r}
ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point() + 
  labs(x = "A", y = "B", title = "C", color = "D")  
```
```{r}
#| fig-cap: "Hi My Name is Suzie"

ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point() + 
  labs(x = "Temp 1", y = "Temp 2", color = "location")  
```
```
```{r}

ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)")  
```

